Sequential and parallel algorithms for multidimensional multiextremal optimization based on the nested scheme of dimensionality reduction

### 2017-10-17 2017/ParOptBook/Chapter5.tex
Chapters 4 and 5 are updated

%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\documentclass[graybox]{svmult}
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%\usepackage{multirow}
%\newcommand{\extr}{extr}
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

@@

As an illustration, let us consider an example of the maximization of a two-dimensional multiextremal function from the class (\ref{eq:3_518}) using the nested optimization scheme. The surface and level curves of the objective function are shown in Fig. 5.1 and in Fig. \ref{fig:5_2}, respectively. The coordinates of the search trials of the nested scheme are shown in Fig. 5.2 also. For this function the global maximum value 13.51 is located at the point (0.603, 0.408). The method AGS (\ref{eq:2_38})--(\ref{eq:2_40}) was employed for solving the one-dimension subproblems (\ref{eq:5_21}).
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_2.pdf}
\caption{Level curves and trial points}
\label{fig:5_2}    
\end{figure}

@@

Then, instead of the family (\ref{eq:5_21}) we can solve the set of nested subproblems 
\begin{displaymath}
\varphi(u_{\sigma(i-1)},z_i)\rightarrow \min,\;z_i\in D_i,
\end{displaymath}
which can be multidimensional with lesser dimensions. Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination of the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.

@@

The structure and the complexity of the projections $\Pi_i(u_{i-1})$  are completely determined by the complexity of the multidimensional feasible domain $Q$. The complexity of the second problem depends on the characteristics of the functions $\varphi^i(u_i)$ . These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined, in turn, by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).

@@

Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous with respect to $u_i\in D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well. Then, at fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
\begin{equation}
\label{eq:5_22}
  \begin{cases}
    \bar{\varphi}(x)\rightarrow\min,\;x\in\bar{Q}\subseteq R^1, \\
    \bar{Q}=\{x\in[a,b]:g(x)\leq 0\},
  \end{cases}
\end{equation}
where the function $g(x)$ is continuous.

The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of intervals
\begin{equation}
\label{eq:5_23}
\bar{Q}=\bigcup_{j=1}^q{[a^j,b^j]},
\end{equation}
in each of which the function $g(x)$ is non-positive.

As an example, let us consider the possible cases of the behavior of a continuous function generating the domains of the non-positivity presented in Fig. \ref{fig:5_3}. The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the points of tangency. The most right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_3.pdf}
\caption{Domains of the non-positivity of a continuous function}
\label{fig:5_3}    
\end{figure}
In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation one can consider within the closed interval $[0,1]$ the function
\begin{displaymath}
g(x)=
  \begin{cases}
    x\sin(\frac{1}{x}), & x>0,\\
    0, & x=0.
  \end{cases}
\end{displaymath}

Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1}$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
\begin{equation}
\label{eq:5_24}
\Pi_i(u_{i-1})=\bigcup_{j=1}^{q_i}{[a_i^j,b_i^j]},
\end{equation}
where, in general, the number of the intervals $q_i$  and their bounds $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
\begin{equation}
\label{eq:5_25}
q_i=q_i(u_{i-1}),\:a_i^j=a_i^j(u_{i-1}),\:b_i^j=b_i^j(u_{i-1}).
\end{equation}

If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable bounds}. In particular, it is possible if all the roots of the functions $G^i(u_{i-1},y_i)$  with regard to the variables $y_i$  can be analytically found.
\begin{example}
\label {exam:5_1}
As the problem to be investigated  the problem (\ref{eq:5_1})--(\ref{eq:5_3}) with 
\begin{equation}
\label{eq:5_26}
  \begin{cases}
    \varphi(y)=y_1^2+y_2^2,\\
   Q=\{y\in D:(y_1-1)^2+(y_2-1)^2-1\leq 0\}, \\
	D=\{y\in R^2:0\leq y_1,y_2\leq 2\}
  \end{cases}
\end{equation}
is considered. The feasible domain is marked with gray in Fig. \ref{fig:5_4}. Also, the level curves of the objective function are shown there. In this problem, the function $G^2(y)=(y_1-1)^2+(y_2-1)^2-1$. It is evident that the function  $G^1(y_1)=\min\{G^2(y_1,y_2):y_2\in [0,\;2]\}=(y_1-1)^2-1$. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_4.pdf}
\caption{Feasible domain and level curves of the objective function}
\label{fig:5_4}    
\end{figure}

According to (\ref{eq:5_16}), the domains where the functions $G^1(y_1)$  for $y_1$  and $G^2(y_1,y_2)$  for $y_2$  are not positive determine the projections $\Pi_1$  and $\Pi_2(y_1)$ , respectively. The domain bounds are the roots of these functions in the interval $[0,2]$. For the function $G^1(y_1)$  these roots are $0$ and $2$. Therefore, $\Pi_1=[0,\:2]$.

@@

Thus, we can write down the bounds of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable bounds (\ref{eq:5_25}).

@@

In the example considered above, we have constructed the one-dimensional search domain bounds  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ . At the same time, one can specify a more visual «geometrical» method for the construction of the projections $\Pi_{i+1}(u_i)$. This method follows from definitions (\ref{eq:5_8}) and (\ref{eq:5_9}) and consists in the construction of the sections of the domain $Q$  by planes $u_i=const$. After that, it is necessary to find the bounds of these sections for the coordinate $y_{i+1}$.

@@

However, finding all roots of a continuous function is a difficult problem, unsolvable analytically, as a rule. In this case, one can try to find these roots numerically. Let us consider, for example, typical problem (\ref{eq:5_22}). In order to find all roots of a continuous function $g(x)$ one can propose  numerical solving  the equivalent optimization problem 
\begin{displaymath}
\left|g(x)\right|\rightarrow \min,x\in [a,b],
\end{displaymath}
which roots of $g(x)$ are the global minima points in. To solve this problem, the characteristical global search algorithms providing the convergence to all global minimum points might be applied.

Another approach to accounting the constraints in the optimization problems (the index method), which does not require to solve any auxiliary problems  is considered in Chapter 4.

@@

The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some particular properties of the constraint functions as well. These properties are the convexity of the functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodality of the constraint functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the concept of convexity.  In general case, monotonous unimodal constraints can generate non-convex domains $Q$.

In the situation of the monotonous unimodality (or of convexity as a particular case) of the constraints $g_j(y),\:1\leq j\leq m$, the one-dimensional search domains are the intervals 
\begin{displaymath}
\Pi_i(u_{i-1})=[a_i^1(u_{i-1}),b_i^1(u_{i-1})],
\end{displaymath}
the bounds of which, however, are not constant but depend on the vector $u_{i-1}$.
\subsection {Properties of the objective functions in one-dimensional subproblems}
\label{subsec:5_2_2}

@@

The function $\varphi^2(y)\equiv \varphi(y)$ , obviously, is Lipschitzian with the constant $L$ with regard to the coordinate $y_2$. However, the function $\varphi^1(y_1)$  does not satisfy the condition (\ref{eq:5_29}). As it has been shown in \cite{5_StrMonRus}, this function satisfies the generalized Lipschitz condition (H{\"o}lder condition) in the metric $\rho(y_1',y_1'')=\sqrt{\left|y_1'-y_1''\right|}$  with the constant $L_1=L(1+\sqrt{2})$ , i.e., the inequality 
\begin{displaymath}
\left|\varphi(y_1')-\varphi(y_1'')\right|\leq L_1\sqrt{\left|y_1'-y_1''\right|}.
\end{displaymath}

Sufficient conditions for the Lipschitz property (\ref{eq:5_29})  of functions $\varphi^i(u_i)$  is established by the following theorem (see \cite{5_StrMonRus}).
\begin{theorem}
\label{theor:5_1}
Assume that the function $\varphi(y)$ satisfies the Lipschitz condition (\ref{eq:5_29})  with the constant $L>0$ in a domain $Q$ from (\ref{eq:5_2}) and the bound pairs (\ref{eq:5_27}) are the piecewise linear functions of the kind 
\begin{equation}
\label{eq:5_30}
a_{i+1}(u_i)=\max_{1\leq\nu \leq p_{i+1}}\{\alpha_i^\nu u_i+A_i^\nu\},
\end{equation}
\begin{equation}
\label{eq:5_31}
b_{i+1}(u_i)=\max_{1\leq\nu \leq r_{i+1}}\{\beta_i^\nu u_i+B_i^\nu\},
\end{equation}
where  $\alpha_i^\nu u_i$ and $\beta_i^\nu u_i$ are the scalar products of the vectors from $R^i$  and $A_i^\nu, B_i^\nu$ are constants. Then the functions $\varphi^i(u_i),u_i\in Q_i,$ satisfy the Lipschitz condition with the constants $L_i,1\leq i\leq N,$ 
where
\begin{displaymath}
L_N=L,\:L_i=L\sum_{j=i}^{N-1}{(1+\lambda_j)},\:1\leq i\leq N-1,
\end{displaymath}
\begin{displaymath}
\lambda_j=\max\left\{\max_{1\leq \nu\leq p_{i+1}}\left\|a_j^\nu\right\|,\max_{1\leq \nu\leq r_{i+1}}\left\|b_j^\nu\right\| \right\}.
\end{displaymath}
\end{theorem}

@@

Note that the representation of the boundy pairs in the form (\ref{eq:5_30}) and (\ref{eq:5_31}) takes place if the feasible domain is a convex polyhedron.

@@

The main conclusion, which follows from the discussion of Lipschitz properties is that the properties of the objective functions of the one-dimensional problems depend essentially not only on the properties of the initial objective function $\varphi(y)$  but also on the shape of the feasible domain $Q$.

@@

One of possible distributions for   $P=6$ parallel processors is presented in Fig. 
\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{figures/figure_5_6.pdf}
\caption{Distribution of processors}
\label{fig:5_6}    
\end{figure}

@@

In order to illustrate the theoretical statements, the results of the numerical solving of different multidimensional multiextremal problems are presented below. Three methods which realize multidimensional optimization using the nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) with combination of information univariate algorithms are considered:
\begin{description} [1)]
\item [1)] {AGSM– algorithm of global search which solves the one-dimensional subproblems inside of the nested scheme by the sequential core information algorithm AGS;}
\item [2)] {PSIM – parallel synchronous information algorithm. It uses AGSIP for solving the one-dimensional subproblems (\ref{eq:5_21});}
\item [3)] {AIPM – asynchronous multidimensional information parallel algorithm applying the asynchronous one-dimensional algorithm proposed in \cite{5_SergGriJCAA} for solving the one-dimensional subproblems.}
\end{description}

@@

This approach allows abstracting from the architecture of the parallel computational system, on which the experiments were carried out and estimate the «pure» efficiency of the algorithm decision rules.

@@

As a measure for the comparison of the relative efficiency of the synchronism and the asynchronism one can introduce the coefficient $u(\pi)=u_A(\pi)/u_S(\pi)$.

@@

In the first experiment, the maximization of 100 two-dimensional functions from the class (\ref{eq:3-58}) has been carried out. All experiments were carried out with the precision in the termination criterion $\epsilon=0.01$  and with the reliability parameter $r=2$  for solving all one-dimensional subproblems (\ref{eq:5_21}). All one-dimensional searches have been started from the same starting points {0.2, 0.3, 0.7, 0.8}.

For the method AGSM, which is a pure sequential one, the average number of trials determined as a result of the maximization of 100 functions of the class (\ref{eq:3-58}) was found to be  $T_1= 415.08$. This value can be considered as a measure for the average search time.

@@

\begin{table}
\caption{Comparison of the parallel algorithms on the two-dimensional test class}
\label {tab:5_1}
\begin{center}
\begin{tabular}{|c|r|r|r|r|r|}
\hline
$q$ & (2,1) & (2,2) & (2,3) & (2,4) & (2,5)   \\
\hline
$\ \ $$K_A\ \ $  &  $\ \ \ $409.46 & $\ \ \ $397.96 & $\ \ \ $419.41 & $\ \ \ $456.00 & $\ \ \ $510.60  \\
\hline
$K_S$  & 413.89 & 409.24 & 424.08 & 458.60 & 514.20  \\
\hline
$T_A$  & 204.73 &  99.49 &  69.90 &  57.00 &  51.06  \\
\hline
$T_S$  & 234.52 & 114.63 &  77.79 &  62.46 &  56.89  \\
\hline
$u_A$  &   2.03 &   4.17 &   5.94 &   7.28 &   8.13  \\  
\hline
$u_S$  &   1.77 &   3.62 &   5.34 &   6.64 &   7.30  \\
\hline
$u$    &   1,15 &   1,15 &   1,11 &   1,10 &   1,23  \\
\hline
\hline
$q$ & (3,1) & (3,2) & (3,3) & (3,4) & (3,5)   \\
\hline
$K_A$  & 427.23 & 429.18 & 432.90 & 472.20 & 529.95  \\
\hline
$K_S$  & 432.40 & 426.16 & 433,83 & 469.44 & 527.90  \\
\hline
$T_A$  & 123.41 &  71.53 &  48.10 &  39.35 &  35.33  \\
\hline
$T_S$  & 174.38 &  84.90 &  55.95 &  44.63 &  40.62  \\
\hline
$u_A$  &   2.91 &   5.80 &   8.63 &  10.55 &  11.75  \\
\hline
$u_S$  &   2.38 &   4.89 &   7.42 &   9.30 &  10.22  \\
\hline
$u$    &   1.22 &   1.19 &   1.16 &   1.13 &   1.15  \\
\hline
\hline
$q$ & (4,1) & (4,2) & (4,3) & (4,4) & (4,5)   \\
\hline
$K_A$  & 459.88 & 447.12 & 469.32 & 494.40 & 565.20  \\
\hline
$K_S$  & 458.09 & 459.38 & 466.44 & 502.08 & 570.70  \\
\hline
$T_A$  & 114.97 &  55.89 &  39.11 &  30.90 &  28.26  \\
\hline
$T_S$  & 144.45 &  71.38 &  46.58 &  36.92 &  33.69  \\
\hline
$u_A$  &   3.61 &   7.43 &  10.61 &  13.43 &  14.69  \\
\hline
$u_S$  &   2.87 &   5.82 &   8.91 &  11.24 &  12.32  \\
\hline
$u$    &   1.26 &   1.28 &   1.19 &   1.19 &   1.19  \\
\hline
\end{tabular}
\end{center}
\end{table}
The next series of the experiments consists in the minimization of the multidimensional function \cite{5_LucidiPiccioni}
\begin{equation}
\label{eq:5_33}

>>>>

%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\documentclass[graybox]{svmult}
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%\usepackage{multirow}
%\newcommand{\extr}{extr}
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

@@

As an illustration, let us consider an example of using the nested optimization scheme for the maximization of a two-dimensional multiextremal function 
\begin{equation}
	\label{eq:5_GriFun}
	\varphi(y)=\left\{\left(\sum _{i=1}^7\sum _{j=1}^7A_{ij} g_{ij} (y)+B_{ij}h_{ij}(y) \right)^2 +\left(\sum _{i=1}^7\sum _{j=1}^7C_{ij}g_{ij}(y)-D_{ij}h_{ij}(y)\right)^2 \right\}^{1/2},
	\end{equation}
	where 
	\begin{equation*}
	y=(y_1 ,y_2 )\in R^2 ,0\le y_s \le 1,s=1,2,
	\end{equation*}
	\begin{equation*}
	g_{ij} (y)=\sin (i\pi y_1 )\sin (j\pi y_2 ),
	\end{equation*}
	\begin{equation*}
	h_{ij} (y)=\cos (i\pi y_1 )\cos (j\pi y_2 )
	\end{equation*}
	 and the coefficients $A_{ij} ,B_{ij} ,C_{ij} ,D_{ij} $ are distributed uniformly over the interval $[-1,1]$.

 The level curves of the objective function are shown  in Fig. \ref{fig:5_2}.  For this function the global maximum value 13.51 is located at the point (0.603, 0.408). Fig. \ref{fig:5_2}contains also the coordinates of the search trials generated by the nested scheme with the information core global search algorithm  applied for solving the one-dimension subproblems (\ref{eq:5_21}).
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_2.pdf}
\caption{Level curves and trial points}
\label{fig:5_2}    
\end{figure}

@@

Then, instead of the family (\ref{eq:5_21}) we can solve the set of nested subproblems 
\begin{displaymath}
\varphi(u_{\sigma(i-1)},z_i)\rightarrow \min,\;z_i\in D_i,
\end{displaymath}
which can be multidimensional with lesser dimensions. Such approach was developed (see \cite{5_BarkGer2014, 5_BarkGerLeb, 5_BarkLeb, 5_SysoyevBarkGerLeb}) in combination with the optimization algorithms based on the Peano space-filling curves for solving the nested multidimensional subproblems.

@@

The structure and the complexity of the projections $\Pi_i(u_{i-1})$  are completely determined by the complexity of the multidimensional feasible domain $Q$. The complexity of the second problem depends on the characteristics of the functions $\varphi^i(u_i)$ . These characteristics are influenced by the properties of the objective function $\varphi(y)$ as well as by the features of the search domain $Q$ defined by constraints (\ref{eq:5_2}) and (\ref{eq:5_3}).

@@

Since the function $G(y)$ is assumed to be continuous in the domain $D$, the functions $G^i(u_i)$ are continuous in $D_i$  from (\ref{eq:5_11}) and, therefore, with regard to $y_i\in [a_i,b_i]$  as well. Then, with fixed $u_{i-1}\in D_{i-1}$  each one-dimensional problem (\ref{eq:5_21}) can be rewritten in a unified form
\begin{equation}
\label{eq:5_22}
  \begin{cases}
    \bar{\varphi}(x)\rightarrow\min,\;x\in\bar{Q}\subseteq R^1, \\
    \bar{Q}=\{x\in[a,b]:g(x)\leq 0\},
  \end{cases}
\end{equation}
where the function $g(x)$ is continuous.

The continuity of the constraint $g(x)$ allows one to state that the feasible domain $\bar{Q}$  can be presented as a system of closed intervals
\begin{equation}
\label{eq:5_23}
\bar{Q}=\bigcup_{j=1}^q{[a^j,b^j]},
\end{equation}
where in each interval the function $g(x)$ is non-positive.

As an example, let us consider the possible cases of the behavior of a continuous function generating the domains of the non-positivity presented in Fig. \ref{fig:5_3}. The corresponding intervals  are marked with  the signs $\times$  on the abscissa axis including the point of tangency. The right-side interval marked with $\times$ corresponds to the case when the function equals to zero over the interval of non-zero length. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_3.pdf}
\caption{Domains of the non-positivity of a continuous function}
\label{fig:5_3}    
\end{figure}
In the system (\ref{eq:5_23}) the number of intervals $q$ can be infinite. As an example of such situation the function
\begin{displaymath}
g(x)=
  \begin{cases}
    x\sin(\frac{1}{x}), & x>0,\\
    0, & x=0.
  \end{cases}
\end{displaymath}
within the closed interval $[0,1]$ can be considered.

Thus, in the case of a continuous function $G(y)$, the projection $\Pi_i(u_{i-1})$ determined as  (\ref{eq:5_16}) is a set of the kind (\ref{eq:5_23}), i.e.,
\begin{equation}
\label{eq:5_24}
\Pi_i(u_{i-1})=\bigcup_{j=1}^{q_i}{[a_i^j,b_i^j]},
\end{equation}
where, in general, the number of the intervals $q_i$  and their boundaries $a_i^j$, $b_i^j$, $1\leq j\leq q_i,$ depend on the vector $u_{i-1}$ , i.e.,
\begin{equation}
\label{eq:5_25}
q_i=q_i(u_{i-1}),\:a_i^j=a_i^j(u_{i-1}),\:b_i^j=b_i^j(u_{i-1}).
\end{equation}

If the domain $Q$  is such that it is possible to obtain the explicit (analytical) expressions for the values $q_i,a_i^j,b_i^j$   as  functions of $u_{i-1})\in Q_{i-1}$  for all $1\leq i\leq N$, then the domain $Q$ is called \textit{the domain with the computable boundaries}. In particular, it is possible if all the roots of the functions $G^i(u_{i-1},y_i)$  with regard to the variables $y_i$  can be analytically found.
\begin{example}
\label {exam:5_1}
As the problem to be investigated  the problem (\ref{eq:5_1})--(\ref{eq:5_3}) with 
\begin{equation}
\label{eq:5_26}
  \begin{cases}
    \varphi(y)=y_1^2+y_2^2,\\
   Q=\{y\in D:(y_1-1)^2+(y_2-1)^2-1\leq 0\}, \\
	D=\{y\in R^2:0\leq y_1,y_2\leq 2\}
  \end{cases}
\end{equation}
is considered. The feasible domain is marked with gray in Fig. \ref{fig:5_4}. Also, the level curves of the objective function are shown there. In this problem, the function $G^2(y)=(y_1-1)^2+(y_2-1)^2-1$. It is evident that the function  $G^1(y_1)=\min\{G^2(y_1,y_2):y_2\in [0,\;2]\}=(y_1-1)^2-1$. 
\begin{figure}[t]
\centering
\includegraphics[width=1.1\linewidth]{figures/figure_5_4.pdf}
\caption{Feasible domain and level curves of the objective function}
\label{fig:5_4}    
\end{figure}

According to (\ref{eq:5_16}), the domains where the functions $G^1(y_1)$  for $y_1$  and $G^2(y_1,y_2)$  for $y_2$  are not positive determine the projections $\Pi_1$  and $\Pi_2(y_1)$ , respectively. The domain boundaries are the roots of these functions in the interval $[0,2]$. For the function $G^1(y_1)$  these roots are $0$ and $2$. Therefore, $\Pi_1=[0,\:2]$.

@@

Thus, we can write down the boundaries of projections (\ref{eq:5_24}) in an explicit form. Thereby, the domain $Q$ from (\ref{eq:5_26}) is a domain with computable boundaries (\ref{eq:5_25}).

@@

In the example considered above, we have constructed the one-dimensional search domain boundaries  analytically by finding the domains of the non-positivity of the corresponding functions $G^i(u_i)$ . At the same time, one can specify a more visual «geometrical» method for the construction of the projections $\Pi_{i+1}(u_i)$. This method follows from definitions (\ref{eq:5_8}) and (\ref{eq:5_9}) and consists in the construction of the sections of the domain $Q$  by planes $u_i=const$. After that, it is necessary to find the boundaries of these sections for the coordinate $y_{i+1}$.

@@

However, finding all roots of a continuous function is a difficult problem, unsolvable analytically, as a rule. In this case, one can try to find these roots numerically. Let us consider, for example, typical problem (\ref{eq:5_22}). In order to find all roots of a continuous function $g(x)$ one can propose  numerical solving  the equivalent optimization problem 
\begin{displaymath}
\left|g(x)\right|\rightarrow \min,x\in [a,b],
\end{displaymath}
which roots of $g(x)$ are the global minima points in. To solve this problem, for example, the characteristical global search algorithms \cite{5_GrishaginSergeyevStrongin} providing the convergence to all global minimum points might be applied.

@@

The optimization domains (\ref{eq:5_2}) can generate projections $\Pi_i(u_{i-1})$  as a single interval for some special classes of the constraint functions as well. These classes are the convex functions $g_j(y),\:1\leq j\leq m$, providing the convexity of the feasible domain $Q$ and the monotonous unimodal functions \cite{5_GrishaginStrongin_EnginCybernetics}, which is more general than the convex ones.  In general case, monotonous unimodal constraints can generate non-convex domains $Q$.

In the situation of the monotonous unimodality (or of convexity as a particular case) of the constraints $g_j(y),\:1\leq j\leq m$, the one-dimensional search domains are the intervals 
\begin{displaymath}
\Pi_i(u_{i-1})=[a_i^1(u_{i-1}),b_i^1(u_{i-1})],
\end{displaymath}
the boundaries of which, however, are not constant but depend on the vector $u_{i-1}$.
\subsection {Properties of the objective functions in one-dimensional subproblems}
\label{subsec:5_2_2}

@@

The function $\varphi^2(y)\equiv \varphi(y)$ , obviously, is Lipschitzian with the constant $L$ for the coordinate $y_2$. However, the function $\varphi^1(y_1)$  does not satisfy the condition (\ref{eq:5_29}). As it has been shown in \cite{5_StrMonRus}, this function satisfies the generalized Lipschitz condition (H{\"o}lder condition) in the metric $\rho(y_1',y_1'')=\sqrt{\left|y_1'-y_1''\right|}$  with the constant $L_1=L(1+\sqrt{2})$ , i.e., the inequality 
\begin{displaymath}
\left|\varphi(y_1')-\varphi(y_1'')\right|\leq L_1\sqrt{\left|y_1'-y_1''\right|}.
\end{displaymath}

Sufficient conditions for the Lipschitz property (\ref{eq:5_29})  of functions $\varphi^i(u_i)$  is established by the following theorem (see \cite{5_StrMonRus}).
\begin{theorem}
\label{theor:5_1}
Assume that the function $\varphi(y)$ satisfies the Lipschitz condition (\ref{eq:5_29})  with the constant $L>0$ in a domain $Q$ from (\ref{eq:5_2}) and the boundary pairs (\ref{eq:5_25}) are the piecewise linear functions of the kind 
\begin{equation}
\label{eq:5_30}
a_i(u_{i-1})=\max_{1\leq\nu \leq p_i}\{\alpha_{i-1}^\nu u_{i-1}+A_{i-1}^\nu\},
\end{equation}
\begin{equation}
\label{eq:5_31}
b_i(u_{i-1})=\max_{1\leq\nu \leq r_i}\{\beta_{i-1}^\nu u_i+B_{i-1}^\nu\},
\end{equation}
where  $\alpha_{i-1}^\nu u_{i-1}$ and $\beta_{i-1}^\nu u_{i-1}$ are the scalar products of the vectors from $R^{i-1}$  and $A_{i-1}^\nu, B_{i-1}^\nu$ are constants. Then the functions $\varphi^i(u_i),u_i\in Q_i,$ satisfy the Lipschitz condition with the constants $L_i,1\leq i\leq N,$ 
where
\begin{displaymath}
L_N=L,\:L_i=L\sum_{j=i}^{N-1}{(1+\lambda_j)},\:1\leq i\leq N-1,
\end{displaymath}
\begin{displaymath}
\lambda_j=\max\left\{\max_{1\leq \nu\leq p_{i+1}}\left\|a_j^\nu\right\|,\max_{1\leq \nu\leq r_{i+1}}\left\|b_j^\nu\right\| \right\}.
\end{displaymath}
\end{theorem}

@@

Note that the representation of the boundary pairs in the form (\ref{eq:5_30}) and (\ref{eq:5_31}) takes place if the feasible domain is a convex polyhedron.

@@

The main conclusion, which follows from the discussion of the Lipschitz property is that the properties of the objective functions of the one-dimensional problems depend essentially not only on the features of the initial objective function $\varphi(y)$  but also on the shape of the feasible domain $Q$.

@@

One of possible distributions for   $P=6$ parallel processors is presented in Fig. \ref{fig:5_6}.
\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{figures/figure_5_6.pdf}
\caption{Distribution of processors}
\label{fig:5_6}    
\end{figure}

@@

In order to illustrate the theoretical statements, the results of the numerical solving of different multidimensional multiextremal problems are presented below. Three methods which realize multidimensional optimization using the nested scheme (\ref{eq:5_18})--(\ref{eq:5_21}) with combination of information univariate algorithms are considered:
\begin{description} [1)]
\item [1)] {AGSM– algorithm of global search which solves the one-dimensional subproblems inside of the nested scheme by the sequential core information global search algorithm \cite{5_StrSergMon2000};}
\item [2)] {PSIM – parallel synchronous information algorithm. It uses the algorithm AGSIP \cite{5_GriKvaMukhStr} for solving the one-dimensional subproblems (\ref{eq:5_21});}
\item [3)] {AIPM – asynchronous multidimensional information parallel algorithm proposed in \cite{5_SergGriJCAA}.}
\end{description}

@@

This approach allows abstracting from the architecture of the parallel computational system, on which the experiments are carried out and estimate the «pure» efficiency of the algorithm decision rules.

@@

As a measure for the comparison of the relative efficiency of the synchronism and the asynchronism the coefficient $u(\pi)=u_A(\pi)/u_S(\pi)$ is introduced.

@@

In the first experiment, the maximization of 100 two-dimensional functions from the class (\ref{eq:5_GriFun}) has been carried out. All experiments were carried out with the precision in the termination criterion $\epsilon=0.01$  and with the reliability parameter $r=2$  for solving all one-dimensional subproblems (\ref{eq:5_21}). All one-dimensional searches have been started from the same starting points {0.2, 0.3, 0.7, 0.8}.

For the method AGSM, which is a pure sequential one, the average number of trials determined as a result of the maximization of 100 functions of the class (\ref{eq:5_GriFun}) was found to be  $T_1= 415.08$. This value can be considered as a measure for the average search time.

@@

\begin{table}
\caption{Comparison of the parallel algorithms on the two-dimensional test class}
\label {tab:5_1}
\begin{center}
\begin{tabular}{|c|r|r|r|r|r|}
\hline
$q$ & (2,1) & (2,2) & (2,3) & (2,4) & (2,5)   \\
\hline
$\ \ $$K_A\ \ $  &  $\ \ \ $409.46 & $\ \ \ $397.96 & $\ \ \ $419.41 & $\ \ \ $456.00 & $\ \ \ $510.60  \\
\hline
$K_S$  & 413.89 & 409.24 & 424.08 & 458.60 & 514.20  \\
\hline
$T_A$  & 204.73 &  99.49 &  69.90 &  57.00 &  51.06  \\
\hline
$T_S$  & 234.52 & 114.63 &  77.79 &  62.46 &  56.89  \\
\hline
$u_A$  &   2.03 &   4.17 &   5.94 &   7.28 &   8.13  \\  
\hline
$u_S$  &   1.77 &   3.62 &   5.34 &   6.64 &   7.30  \\
\hline
$u$    &   1,15 &   1,15 &   1,11 &   1,10 &   1,23  \\
\hline
\hline
$q$ & (3,1) & (3,2) & (3,3) & (3,4) & (3,5)   \\
\hline
$K_A$  & 427.23 & 429.18 & 432.90 & 472.20 & 529.95  \\
\hline
$K_S$  & 432.40 & 426.16 & 433,83 & 469.44 & 527.90  \\
\hline
$T_A$  & 123.41 &  71.53 &  48.10 &  39.35 &  35.33  \\
\hline
$T_S$  & 174.38 &  84.90 &  55.95 &  44.63 &  40.62  \\
\hline
$u_A$  &   2.91 &   5.80 &   8.63 &  10.55 &  11.75  \\
\hline
$u_S$  &   2.38 &   4.89 &   7.42 &   9.30 &  10.22  \\
\hline
$u$    &   1.22 &   1.19 &   1.16 &   1.13 &   1.15  \\
\hline
\hline
$q$ & (4,1) & (4,2) & (4,3) & (4,4) & (4,5)   \\
\hline
$K_A$  & 459.88 & 447.12 & 469.32 & 494.40 & 565.20  \\
\hline
$K_S$  & 458.09 & 459.38 & 466.44 & 502.08 & 570.70  \\
\hline
$T_A$  & 114.97 &  55.89 &  39.11 &  30.90 &  28.26  \\
\hline
$T_S$  & 144.45 &  71.38 &  46.58 &  36.92 &  33.69  \\
\hline
$u_A$  &   3.61 &   7.43 &  10.61 &  13.43 &  14.69  \\
\hline
$u_S$  &   2.87 &   5.82 &   8.91 &  11.24 &  12.32  \\
\hline
$u$    &   1.26 &   1.28 &   1.19 &   1.19 &   1.19  \\
\hline
\end{tabular}
\end{center}
\end{table}
